  ---
title: "Script for Figures from - (Mora Van Cauwelaert et al., 2022)"

output: html_notebook
---


This script generates Fig.3, 4, 5, 6, S1.3, S1.4, S1.5. and send them to the **output_figure** 

Load Packages and colors

```{r}
mycols <- c("#021128","#1b4a64", "#3585a0", "#759580", "#c78f34", "#fd9706","#fdb81c","#fbdb30")
mycols3a <-c("#021128", "#fd9706", "#3585a0")
mycols3b <-c("#1b4a64", "#fdb81c", "#759580")
library(ggplot2)
library(patchwork)
library(tidyverse)

```

# Part 1 (Figures 3 and S1.3 One plant-level simulation) 

## 1.1. Fig 3.AB

```{r echo=TRUE}

datosFig_local<- read.csv("./onePlant_data/dataIndividualPlant.csv")
datosFig_local$Ieq <- datosFig_local$Imax1-datosFig_local$Imax1.Ieq

#quitar los ultmos valores
datosFig_local <-  datosFig_local[datosFig_local$g!= 0.2,]
datosFig_local$I0 <- datosFig_local$I0*100

datosFig_local <- datosFig_local %>%
  filter(g<0.14)

#todo esto ya no va  estar
#cortar la simulacion a 300 (osea todos los valores mayores a 300, asumimo que fueron 300..) VERIF
#datosFig_local$tiempoImax1[datosFig_local$tiempoImax1 >300] <- 300 


Fig3A <- datosFig_local  %>% 
  filter(!(I0==2.3)) %>% 
  filter(!(I0==6.7)) %>% 
  ggplot(aes(g, Imax1))+  #change to Imax.Ieq for supp figure S1.2
  theme_bw()+
  geom_point(aes(color=as.factor(I0)), size=2) +  #corregir esto
  scale_color_manual(values= mycols)+
  theme(text = element_text(size=15))+
  theme(legend.position = "none")+
  labs(color= "Initial proportion of infected leaves", x= "Leaf-fall rate (\u03B3)", y= "MATI") 

Fig3B <- datosFig_local  %>% 
  filter(!(I0==2.3)) %>% 
  filter(!(I0==6.7)) %>% 
  ggplot(aes(g, tiempoImax1))+
  theme_bw()+
  theme(text = element_text(size=15))+
  theme(legend.position = "none")+
  geom_point(aes(color=as.factor(I0)), size=2) +  #corregir esto
  scale_color_manual(values= mycols)+
  labs(color= "Initial % of infected leaves", x= "Leaf-fall rate (\u03B3)", y= "Time to MATI") 
  

Fig3AB <- Fig3A+Fig3B

ggsave("./output_figures/fig3_AB.png", height = 4, width = 10)

```


## 1.2. Fig 3.CD

```{r echo=TRUE}
d_ex_plant<- read.csv("./onePlant_data/examplesIndividualPlant.csv")

Nth.delete<-function(dataframe, n)dataframe[-(seq(n,to=nrow(dataframe),by=n)),] #funcion bajada para borrar lineas
d_ex_plant<- Nth.delete(d_ex_plant, 2)
d_ex_plant<- Nth.delete(d_ex_plant, 2)

library(dplyr)

#primerFiltro
parEj1<- filter(d_ex_plant, g == 0.015)
parEj2<- filter(d_ex_plant, I0== 0.01)

parEj1 <- parEj1 %>%
  filter(X.1<30001)
parEj2 <- parEj2 %>%
  filter(X.1<30001)


Fig3C <- ggplot(parEj1)+
  geom_line(aes(x=X.1*0.01, y=S, linetype=as.factor(I0)), color="darkgreen", size=1) +
  geom_line(aes(x=X.1*0.01, y=I, linetype=as.factor(I0)), color="darkred", size=1)+
  theme_bw()+
  theme(text = element_text(size=15))+
  theme(legend.position = "none")+
  labs(linetype= "I0", x= "Time (days)", y= "S and I (densities)")


Fig3D <- ggplot(parEj2)+
  geom_line(aes(x=X.1*0.01, y=S, linetype=as.factor(g)), color="darkgreen", size=1) +
  geom_line(aes(x=X.1*0.01, y=I, linetype=as.factor(g)), color="darkred", size=1)+
  theme_bw()+
  theme(text = element_text(size=15))+
  theme(legend.position = "none")+
  labs(linetype= "\u03B3", x= "Time (days)", y= "S and I (densities)")

    
  Fig3CD <- Fig3C+Fig3D
  #save figure
ggsave("./output_figures/fig3_CD.png", height = 4, width = 10)
```

# Part 2 (figures 4 (general MATI and time to MATI vs m), S1.4 (MATI vs <D>) and  5 (time series))

## 2. 1. Load the data frame and modify it only to plot the required values

```{r}

datosFig_Espacio <- read.csv("./lattice_data/dataGeneralAverages.csv")

#add an ID per simulations  (990 IDs)
datosFig_Espacio$ID <- rep(1:990, each= 602)  # 602 is the number of steps per simulations after the filter
###transform steps to days 
datosFig_Espacio$X <- datosFig_Espacio$X*0.01 

#remove the middle value of Iini and filter the low value of g
datosFig_Espacio <- datosFig_Espacio %>% 
  filter(!(Iini==0.010)) %>%  
  filter(g == 0.015)  
#%>%  
 # filter(g== 0.056) 
```

## 2.2. Database modifications for figures 4 and 5

```{r}

#obtain the  maximum values of each serie, according to the ID
datosMax<- datosFig_Espacio %>% group_by(g,m, Iini, sim, patron, Dc, ID) %>%
  summarise_all(max)

#subselct in the general data frame rows that contain these maximum values. This procedure works since the values have mutiple decimals, and we don't have to simulations with the exact same value

dgInf <- datosFig_Espacio[datosFig_Espacio$I_prom %in% datosMax$I_prom, ]
dgInf<- dgInf %>% select(!c(S_prom,S_sd,X_prom,X_sd, I_sd))

#we merge the simulatons per planting pattern to summarise the random simulations and have an average and a standard deviation, only for that planting pattern
dgInf_ave <- dgInf  %>% group_by(g,m, Iini, patron) %>%  
  summarise(meanDc= mean(Dc), sdDc= sd(Dc), meanImax=mean(I_prom), sdImax= sd(I_prom),meanTime=mean(X), sdTime= sd(X), ID=mean(ID) ) 

```


## 2.3. Procedure to define the time lag and the growth per se from each averaged run. 

We estimate the day when the slope was higher than 0.001, this is, an increase of 0.1% of infection per day
```{r}

#We estimate the slope for each run
y2 <- datosFig_Espacio$I_prom[2:length(datosFig_Espacio$I_prom)]
y1 <- datosFig_Espacio$I_prom[1:length(datosFig_Espacio$I_prom)-1]

change <- y2-y1
change <- append(change, 0, 0)
datosFig_Espacio$slope <- change/0.5 # 0.5 is the time step between each point

#we now remove the values of the slopes for the first value or each simulation (since it does not make sense)
datosFig_Espacio$slope[datosFig_Espacio$X == 0.49] <- 0

#We group by Id and filter the first value where the slope is higher than 0.001

datosSLOPE <- datosFig_Espacio %>% 
  group_by(ID) %>%
  filter(slope>0.001) %>%
  filter(row_number()==1)

#we average in order to have the mean values for the random pattern

datosSLOPE$TimeMax <- dgInf$X
datosSLOPE$TimeGrow <- datosSLOPE$TimeMax- datosSLOPE$X 

datosSLOPE_ave <- datosSLOPE   %>% group_by(g,m, Iini, patron) %>%  
  summarise(meanTime= mean(TimeMax), sdTime= sd(TimeMax), meanTimeGrow= mean(TimeGrow), sdTimeGrow= sd(TimeGrow), meanTimelag= mean(X), sdTimelag= sd(X), ID=mean(ID)) 

#we merge the new values to the previous data frame, by ID. 


# this code generates the necessary csv for table S3.
write.csv(datosSLOPE_ave, "./output_figures/TimeLag_TimeGrow.csv")
```


## 2.4. Figure 4 (MATI and time to MATI vs m)

```{r}

Fig4_A<- dgInf_ave %>% 
 ggplot(aes(x=log10(m)))+
  geom_line(aes(y = meanImax, col=as.character(Iini)), size=0.5) +
  geom_point(aes(y = meanImax, col=as.character(Iini)), size= 2) +
  geom_errorbar(aes(ymin = meanImax - sdImax, ymax = meanImax + sdImax, col=as.character(Iini)),alpha=0.75) +
  facet_wrap(~ patron, ncol = 1) +
  scale_color_manual(values= mycols3a)+
  theme_bw()+
  theme(strip.background = element_rect(color="white", fill="white", size=1.5))+
  theme(text = element_text(size=15))+
  theme(legend.position = "none")+
  labs(x= "Difusion rate [log(m)]", y= "MATI", col= "Initial proportion of infected trees")

#Save the figure

#gather for ggplot viusalization
dgInf_bars <- gather(datosSLOPE_ave, "classTime", "value", c(7,9))
dgInf_bars$Iini_classTime <- paste(dgInf_bars$Iini, dgInf_bars$classTime)

#this figure is not the final fig 4 figure (we editted with inkscape using this figure)
Fig4_B<- dgInf_bars  %>%
  ggplot(aes(x = as.numeric(interaction(log10(m),Iini)), y= value, group=Iini_classTime, fill= Iini_classTime))+
  geom_bar(stat= "identity", width = 0.3)+
  geom_errorbar(aes(ymin = meanTime - sdTime, ymax = meanTime + sdTime),col="black",alpha=0.75, width = 0.2) +
  facet_wrap(~ patron, ncol = 1) +
  scale_fill_manual(values= mycols[c(1,2,6, 7)])+
  theme_bw()+
  theme(strip.background = element_rect(color="white", fill="white", size=1.5))+
  theme(text = element_text(size=15))+
  theme(legend.position = "none")+
  scale_x_continuous(breaks=c(-3,-2.5,-2, -1.5, -1), labels =c("-3","-2.5","-2", "-1.5", "-1"))+ ##VER
  labs(x= "Difusion rate [log(m)]", y= "Time to MATI (Days)", col= "Initial infected trees (%)")

#save the figure 
Fig4 = Fig4_A + Fig4_B

ggsave("./output_figures/fig4.pdf", height = 10, width = 8)
```


## 2.5. Figure S1.4 (MATI and time to MATI vs <D>)

```{r}
FigS1_4_MATIvsD<- dgInf_ave%>%
  filter(Iini == 0.001)%>%
  #filter(!(m==0.1)) %>%
  #filter(!(m==0.005)) %>%
  #filter(!(m==0.05)) %>%
  #g == 0.056 
  #  filter(!(IresPro==5)) %>% 
  #filter(!(meanTime>399))%>%  #esnayo
  ggplot(aes(x=meanDc, y= meanImax))+
  geom_point(aes(y = meanImax, group= as.factor(m)), size=2) +
  geom_errorbar(aes(ymin = meanImax - sdImax, ymax = meanImax + sdImax, col= as.factor(Iini)),alpha=0.75) +
  geom_errorbarh(aes(xmin = meanDc - sdDc, xmax = meanDc + sdDc, col=as.factor(Iini)),alpha=0.75)+
  geom_line(aes(y = meanImax, group= as.factor(m), linetype=as.factor(m)), size=0.5) +
  #facet_wrap(~ Iini, ncol = 1) +
  scale_color_manual(values= mycols[c(1,6,2,7)])+
  theme_bw()+
  theme(strip.background = element_rect(color="white", fill="white", size=1.5))+
  theme(text = element_text(size=15))+
  theme(legend.position = "none")+
  labs(x= "Pattern Index (D)", y= "MATI", col= "Initial proportion of infected trees")


FigS1_4_timeMATIvsD<- dgInf_ave%>%
  filter(Iini == 0.001)%>%
  #filter(!(m==0.1)) %>%
  #filter(!(m==0.005)) %>%
  #filter(!(m==0.05)) %>%
  #g == 0.056 
  #  filter(!(IresPro==5)) %>% 
  #filter(!(meanTime>399))%>%  #esnayo
  ggplot(aes(x=meanDc, y= meanTime))+
  geom_point(aes(y = meanTime, group= as.factor(m)), size=2) +
  geom_errorbar(aes(ymin = meanTime - sdTime, ymax = meanTime + sdTime, col= as.factor(Iini)),alpha=0.75) +
  geom_errorbarh(aes(xmin = meanDc - sdDc, xmax = meanDc + sdDc, col=as.factor(Iini)),alpha=0.75)+
  geom_line(aes(y = meanTime, group= as.factor(m), linetype=as.factor(m)), size=0.5) +
  #facet_wrap(~ Iini, ncol = 1) +
  scale_color_manual(values= mycols[c(1,6,2,7)])+
  theme_bw()+
  theme(strip.background = element_rect(color="white", fill="white", size=1.5))+
  theme(text = element_text(size=15))+
  theme(legend.position = "none")+
  labs(x= "Pattern Index (D)", y= "Time to MATI (days)", col= "Initial proportion of infected trees")

FigS1_4 = FigS1_4_MATIvsD + FigS1_4_timeMATIvsD

ggsave("./output_figures/figS1_4.png", height = 4, width = 8)

```



## 2.6. Time series of figure 5.

We know subselect from the general dataframe of averages,  some specific values of m, to plot the time series in figure 5. 

```{r}
datosFig_Espacio$m[datosFig_Espacio$m == "0.001"] <- " Low diffusion rate" # le pongo un espacio de más para el orden final de la figura
datosFig_Espacio$m[datosFig_Espacio$m == "0.01"] <- "High diffusion rate"


Fig5_timeSeries <- datosFig_Espacio %>%
  filter(!(m==0.1)) %>%
  filter(!(m==0.005)) %>%
  filter(!(m==0.05)) %>%
  filter(sim ==1 | sim==2) %>% # we take each simulation and 1 simulation for the random
  ggplot(aes(x=X))+
  geom_line(aes(y=I_prom, linetype=as.factor(Iini)), color="darkred")+
  facet_wrap(~patron*m, ncol = 4, labeller = label_wrap_gen(multi_line=TRUE)) +
  theme_bw()+
  theme(text = element_text(size=15))+
  theme(panel.spacing.y = unit(6, "lines"))+
  theme(panel.spacing.x = unit(1, "lines"))+
  theme(legend.position = "none")+
  theme(strip.background = element_rect(color="white", fill="white", size=1.5))+
  scale_fill_manual(values= mycols3a)+
  labs(linetype= "I0", x= "Time (Days)", y= "Average tree infection (Infected leaves density)")

#save figure
Fig5_timeSeries
ggsave("./output_figures/fig5.png", height = 8, width = 10)
```



# Part 3. Figure 6 (histograms)
## 3.1  Load and modify data base

```{r}
datosPicosPlanta <-read.csv("./lattice_data/dataPeaks2.csv", header = TRUE)
```



```{r}
datosHisto <- datosPicosPlanta %>%
   filter(TiempoImax<302) %>%
  filter(g==0.015) %>%
  #filter(g==0.056) %>%  ##use only for SUPP
  
   filter(!(tasaMig==0.005)) %>%
   filter(!(tasaMig==0.05)) %>%
   filter(!(tasaMig==0.1)) %>%
   filter(!(Ini ==0.01))%>%
  filter(Sim ==1 | Sim==2)
```


```{r}

datosHisto$numCat <- as.integer(datosHisto$TiempoImax/14)
datosHisto<- datosHisto %>%  unite("mig_Ini", tasaMig, Ini, remove = FALSE)

datosHisto$tasaMig <- as.character(datosHisto$tasaMig)
datosHisto$tasaMig[datosHisto$tasaMig == "0.001"] <- " Low diffusion rate" 
datosHisto$tasaMig[datosHisto$tasaMig == "0.01"] <- "High diffusion rate"

# grooup by pattern to count the number of trees that reach the max in the same cat
datosTotal <- datosHisto %>% 
  filter(Imax>0.7) %>%
  group_by(Sim, Patron, tasaMig, g, Dc, Ini, ID, numCat) %>%   #aqui quite el tiempoImax
  summarise(contadorAgg = sum(contador), ImaxPro = mean(Imax))  #aqui hice un cambio

# for each ID, make the cumulative sum
datosTotal <- datosTotal %>% 
  group_by(ID) %>%
  mutate(csum = cumsum(contadorAgg)/50)  # el 

#visualisation
datosTotal$Ini <- as.character(datosTotal$Ini)
datosTotal$Ini[datosTotal$Ini== "0.001"] <- " Low initial infection (I0)" # le pongo un espacio de más para el orden final de la figura
datosTotal$Ini[datosTotal$Ini == "0.1"] <- "High initial infection (I0)"
datosTotal<- datosTotal %>%  unite("mig_Ini", tasaMig, Ini, remove = FALSE)
```

## 3.2 Fig 6- Histogram
```{r}
Fig6_histoAcc <-   datosTotal%>% 
  ggplot(aes(x= numCat*14))+
#  geom_point(aes(color = mig_Ini), size = 2)+
  geom_col(width=14, aes(y= contadorAgg/50, fill = mig_Ini)) +#geom_errorbar(aes(ymin=avTrees-sdTrees, ymax=avTrees+sdTrees, color=mig_Ini), size=1)+
  geom_line(aes(y= csum, group= mig_Ini), size = 0.5, color = "black")+
  geom_point(aes(y = csum, color = mig_Ini), size =1.5)+
  theme_bw()+
  theme(text = element_text(size=15))+
  theme(panel.spacing.y = unit(1, "lines"))+
  theme(legend.position = "none")+
  theme(strip.background = element_rect(color="white", fill="white", size=1.5))+
  scale_color_manual(values= mycols[c(1,6,3,8)])+
  scale_fill_manual(values= mycols[c(1,6,3,8)])+
  facet_grid(Patron~Ini*tasaMig)+
  labs(linetype= "I0", x= "Time (Days)", y= "Proportion of trees that reach a high maximum infection")

Fig6_histoAcc

ggsave("./output_figures/Fig6.png", height = 8, width = 10)

```


Supplemantary Fig S1.5 Maximum individual tree infections
 
```{r}
FigS1_5<-   datosHisto%>% 
  #filter(Imax >0.5)%>% 
  ggplot(aes(x= TiempoImax, y= Imax))+
#  geom_point(aes(color = mig_Ini), size = 2)+
  geom_point(aes(fill = mig_Ini), size =3, colour="black",pch=21)+
  theme_bw()+
  theme(text = element_text(size=15))+
  theme(panel.spacing.y = unit(1, "lines"))+
  theme(legend.position = "none")+
  theme(strip.background = element_rect(color="white", fill="white", size=1.5))+
  scale_color_manual(values= mycols[c(1,6,3,8)])+
  scale_fill_manual(values= mycols[c(1,6,3,8)])+
  facet_grid(Patron~Ini*tasaMig)+
  labs(linetype= "I0", x= "Time (Days)", y= "Maximum individual tree infection")


ggsave("./output_figures/FigS1_5.png", height = 8, width = 10)

```





